---
title: "Penguins ANOVA"
author: "Yingfei Jiang"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in penguins.csv data

```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>%
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
             )
         ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

```

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length))+
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name)

hists

qq <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq()+
  facet_wrap(~common_name)

qq



# Levene's test for equal variances (> 2 groups ... if you have two groups, use the F-test for equal variance)

# We'll use the fuction leveneTest from the 'car' package

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)

penguin_levene

# We retain the null hypothesis of equal variance (p = 0.06)

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances



```

#3. One-way ANOVA

Single factor: common_name (species)
Number of levels in the factor: 3 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo and Adelie penguins

```{r}

#H0: mean flipper lengths across all species are equal
#HA: At LEAST two means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# At least two samples were taken from populations with different means.
# There is a significant difference in flipper length based on species

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

```

